Identification of protoneutron star g-modes in gravitational-wave data

### 2020-07-01 methods.tex
more updates

\section{Methods description}
In this section, we outline a strategy for estimating the time evolution of the
ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the PNS and its squared radius
(in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
oscillation mode in the gravitational wave detector data.
An integral part of this strategy is the universal relationships that relate the
characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
and the radius of the PNS, the shock radius and the total mass inside the shock as
demonstrated in \cite{Torres:2019}.

@@

To develop the method we considered the gravitational wave signal
{\tt s20-gw-10kpc} described in Section \ref{sec:simulation}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode and we have
developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
taking into account that it is monotonically increasing as time goes.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

@@

We identify the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use the equation (\ref{eq:universal}) to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. The black points are the true ratio values
computed using the mass and radius values obtained from the simulation code.

@@

%In this case without any noise, the coverage of our 95\% confidence band is xx\%.

In the following simulation study we explore how accurately we can estimate the parameters
when the gravitational wave signal is embedded in noise.
For that purpose, we inject the gravitational wave signal into  simulated Advanced LIGO noise
using the noise power spectral density \textcolor{red}{insert formula} for varying
SNRs, respectively distances to the source. We estimate the coverage probability of the 95\%
confidence band by calculating the proportion of times that the true ratio lies outside one
of the pointwise 95\% confidence intervals.
These coverage probabilities together for varying SNRs are given in Table
\textcolor{red}{insert Table} and displayed in the form of boxplots in Figure
\textcolor{red}{insert Figure}

>>>>

\section{Methods description}
\label{methods}

In this section, we outline a strategy for estimating the time evolution of the
ratio $r=M_{\rm PNS}/R_{\rm PNS}^2$ of the mass of the PNS and its squared radius
(in units of solar mass and km) from the observation of the $\mbox{}^2g_2$
oscillation mode in the gravitational wave detector data.
An integral part of this strategy is the universal relations that relate the
characteristic frequency of the PNS oscillation $f$, $g$ and $p$ modes with the mass
and the radius of the PNS, the shock radius and the total mass inside the shock as
demonstrated in \cite{Torres:2019}.

@@

To develop the method we considered the gravitational wave signal
{\tt s20-gw-10kpc} described in Section \ref{sec:simulations}, originally
sampled at 16384 Hz but resampled at 4096 Hz.
A spectrogram of this signal is shown in Figure \ref{fig:spectrogram} based on
autoregressive estimates of the local spectra for successive time intervals of 
\textcolor{red}{length 200} with a \textcolor{red}{ 90\%} overlap.
The dominant emission mode corresponds to the $\mbox{}^2 g_2$-mode and we have
developed a time-frequency method to track the ridge $m(t)$ in the spectrogram,
taking into account that it is monotonically increasing as time goes.
Starting from either the left- or right-most column of the time-frequency matrix
we identify and trace the sequence of amplitude peaks within a certain frequency
band given the monotonicity constraint. Appendix \ref{app:gmode} is providing more
details on the reconstruction of the $g$ mode ridge.

@@

We collect the instantaneous frequency $f(t_i)$ corresponding to the ridge $m(t_i)$ for
the midpoint $t_i$ of each local time interval of the spectrogram and interpolating $f(t)$
for values in between the $t_i$. We then use equation (\ref{eq:universal}) to obtain
estimates of the time evolution of the ratio together with 95\% confidence intervals.
An exemple is given in Figure \ref{fig:ratio} where the red points are the point estimates and
the grey bands represent 95\% confidence bands. The black points are the true ratio values
computed using the mass and radius values obtained from the simulation code.

@@

In this case, for a GW signal without any noise, the coverage of our 95\% confidence band is 94\%.
In the next section we explore how accurately we can estimate $r(t)$ when the gravitational wave
signal is embedded in noise.
